---
title           : "Preregistration: Effects of white-nose syndrome (WNS) on bat community structure in Cypress Hills, Saskatchewan, Canada."
shorttitle      : "PreReg: Bat Community Structure in Cypress"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Sophia Fan
    affiliation : 1
  - name        : Keerthi Krutha
    affiliation : 3
  - name        : Javad Meghrazi
    affiliation : 1
  - name        : Tanya Strydom
    affiliation : 2
  - name        : Hannah Wilson
    affiliation : 3
      
affiliation:
    
  - id          : 1
    institution : University of British Columbia
  - id          : 2
    institution : University of Regina
  - id          : 3
    institution : Université de Montréal

output: prereg::cos_prereg
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

# Study Information

## Title

`r rmarkdown::metadata$title`

## Description

White-nose syndrome (WNS) is an infectious fungal disease caused by
*Pseudogymnoascus destructans* targeting hibernating bats. First
detected in New York City in the winter of 2006, it has since affected
and killed more than 6 million hibernating bats in eastern North America
@jachowski2014. Notably, the disease has had a pronounced effect on the
little brown bat, *Myotis lucifugus*, where it's predicted that regional
population collapse and extirpation could occur as soon as within the
next 16 years @frick2010a. Other bat species responded variably: some
studies show that silver-haired bat (*Lasionycteris noctivagans*)
detection rates increased following WNS [@jachowski2014a; @nocera2019],
while results for big brown bats (*Eptesicus fuscus*) are mixed, but
most studies seem to show that their numbers increased following
competitive relaxation from other declining bat species
[@jachowski2014b; @francl2012; @morningstar2019; @nocera2019a]. Because
WNS affects bat species differentially, community-level monitoring
studies are needed to fully appreciate and mitigate consequences for
community structure and ecosystem function @hoyt2021. In 2021, the first
instance of *P. destructans* with no bat mortality was reported in
Saskatchewan, Canada (Wilson pers. comm., Lausen-WCS Canada pers.
comm.), indicating westward disease spread into central Canada. This
study aims to inform local bat conservation efforts by compiling and
comparing bat community data from before and after the introduction of
*P. destructans* in the Cypress Hills, in addition to constructing
population projection models to predict species survivorship into the
future.

## Hypotheses

We expect to see white-nose syndrome affect bat species differently in
the Cypress Hill region of Saskatchewan, Canada, in line with similar
trends found in affected eastern North American bat populations (Figure
1). Specifically:

**Directional**: We expect to see pronounced population **declines** in
little brown bats (*Myotis lucifugus*) after *P. destructans*
introduction.

**Directional**: We expect to see population **increases** in
silver-haired bats (*Lasionycteris noctivagans*) and big brown bats
(*Eptesicus fuscus*) after *P. destructans* introduction due to reduced
interspecific competition.

# Design Plan

## Study type

**Observational Study**. Data is collected from study subjects that are
not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.

## Blinding

No blinding is involved in this study.

## Study design

**Design:** Case-control study design [@mann2003]. The study group will
be the bat community in Saskatchewan. Our control group would be the bat
community in the Cypress Hills before the detection of *P. destructans.*

**Assumption(s):** No two same individual bats were sampled before and
after *P. destructans* detection. Proportion of each species captured
are equal to their proportion in nature. Other factors that may impact
species decline are minimal.

## Randomization

Not applicable

# Sampling Plan

## Existing data

**Registration prior to analysis of the data**. As of the date of
submission, the data exist and you have accessed it, though no analysis
has been conducted related to the research plan (including calculation
of summary statistics). A common situation for this scenario when a
large dataset exists that is used for many different studies over time,
or when a data set is randomly split into a sample for exploratory
analyses, and the other section of data is reserved for later
confirmatory data analysis.

## Explanation of existing data

Our dataset will be from 2010-2030, and the dataset from 2010-2021
already exists. Our access to the dataset has been limited because, all
team members except HW come from other departments and laboratories with
no prior studies conducted on bats. HW is a new student to the bat
research group to the University of Regina and started in September
2021. The 2010-2021 dataset we plan to use for our study has been
collected by multiple trained researchers who are not connected with our
project. HW has seen the dataset with data from 2019 but has not
conducted any form of statistical analysis with it. Also, as authors of
this study, none of us are aware of any patterns or summary statistics
relating to the effect of a fungal disease on bats of this dataset even
though there are publications/results from previous studies conducted by
other researchers. We plan to only conduct any detailed analysis with
the data only after 2030 once all data collection is complete.

## Data collection procedures

**Study site:** West Block, Cypress Hills Provincial Park, Saskatchewan,
Canada.

**Study locations:** Previously established capture locations for bats
along Battle Creek, established by the University of Regina (Wilson
pers. comm.).

**Sampling period:** Mid-June to mid-August every year. Samples will be
collected until August 2030. Data points from this same period will be
used from records prior to *P. destructans* incidence starting from
August 2010.

**Data collection methods:** We will set up 1-3 mist nets at one site
each night. Nets will be set up across Battle Creek. Nets will be at
least 1m tall, and wide enough the cover the width of the creek. We will
raise nets for three hours after last light and check the nets every
10min in the first hour after last light and every 15min in the
subsequent two hours. We will capture and handle bats with
nitrile-gloves. Once a bat is caught it will be moved away from the net
in order to avoid attracting other individuals to the captured bat's
distress call. We will record the age, sex, mass and species of the bat.

In order to attribute the possible difference in bat populations due to
the fungal disease we will collect samples of *P. destructans* by
swabbing the wings and muzzle of the bat following @zukal2016. We will
analyse swabs using qPCR techniques @zukal2016 and record whether the
individual bat has contracted *P. destructans*.

We will also measure ambient temperature, cloud cover, and wind speed
every night at each sampling site. In case of rain or if wind speed is
above 20km/s, we will stop sampling. When recording a capture, we will
record the date of sampling in the ISO format along with a three-letter
code of the sampling location, the genus and species of bat being
captured and initials of the personnel collecting the capture.

## Sample size

Our target sample size is 2000 individuals. However, we do not have a
definitive pre-determined count and cannot control the number of bats we
capture each year.

## Sample size rationale

We do not have control over how many bats will be captured in a night.
From anecdotal knowledge and previous experience we estimate that we
will to capture anywhere between 100-200 bats per summer. Therefore,
after twenty years of data collection we will have collected around 2000
bats.

## Stopping rule

The station where researchers live during the field season is owned by
the University of Regina, who use the station for an undergraduate class
during the last week of August. The field station is closed after that,
so data collection will stop mid-August.

The project overall will end in 2030. *P. destructans* was first
detected in Saskatchewan in 2021. Ending the project 10 years after the
fungus was first detected allows us to observe bat community structure
at the start of the disease, directly after it should have killed the
largest number of bats, and ten years later when it should have become
endemic to the population.

# Variables

## Manipulated variables

'Raw' bat species abundance records will be transformed to a
proportional abundance which will be used in the modelling process.

## Measured variables

We are going to be measuring the population size of the bat species in
the Cypress Hills before and after the detection of *P. destructans*. We
will measure this by counting how many bats of each species we catch in
mist nets each year and recording instances of infection.

## Indices

Not applicable.

# Analysis Plan


### Are there significant changes in population size for bat species in Cypress Hills before and after the introduction of *P. destructans*?

Unpaired student t-tests will be used to compare population size
estimates for each species before and after infection. Specifically,
between 2021 (summer before detection) and 2023 (two winters later), as
well as at any increments of interest up to 2030. The test should be one
tail. In addition, power analysis should be conducted using `pwr`
package in R. We expect to see significant differences in population
size before and after infection.

### Predictive model of bat community composition change

See Figure 1 for how we expect model output to potentially look

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("data/sim_data.R")
```

```{r plot, echo=FALSE, fig.cap="Expected changes in bat species frequencies after the introduction of white nose syndrome in 2019 the Cypress hill region of Saskatchewan Province.", message=FALSE, warning=FALSE}
p
```


To investigate effect of WNS on bat community as a whole instead of
investigating species individually, we have also developed a
mathematical framework to compare Cypress Hills data with existing
eastern North American data. The following equations aim to describe
changes in the size of populations of different bat species. The aim of
the model is to simulate bat population data for Cypress Hills following
infection. Some coefficients should be estimated based on Saskatchewan
bat community data from 2010-2021 (growth rate, carrying capacity and
competition coefficients) and some from previous studies in east coast
(transmission rate) populations and then predictions can be made for
Saskatchewan bat community. $$
N_{i,t+1} = N_{i,t}r_i (1+\frac{K_i-N_{i,t}- \sum_{i = 1}^{n}{\alpha_{i,j} N_{j,t}}}{K})-I_{i,t}
$$ Where: $N_{i,t}$ is the population size of species i at time t.

$r_i$ is the intrinsic growth rate of species i.

$K_i$ is the carrying capacity for species i.

$\alpha_{i,j}$ is the coefficient for competition effect of species j on
species i.

$I_{i,t}$ is the number of individuals infected with WNS at time t. It's
assumed that all of these individuals will die before next time point.
This parameter is calculated as follows: $$
I_{i,t+1}=d_iI_{i,t}\frac{N_{i,t}}{k_i}
$$ Where $d_i$ is the transmission rate of WNS in species i. It's
assumed that transmission of disease is density dependent.

Estimation of coefficients: $r$, $K$ and $\alpha$: These should be
calculated based on the data gathered up to 2021 using a genetic
algorithm.

$d$: Will depend on the objective of study: 1) If we are going to make
predictions for future dynamics of populations, previous studies should
be used to estimate this parameter. It should be considered that $d$
likely won't stay constant over time and evolutionary dynamics of host
and parasite would affect it. 2) If we are going to estimate $d$ in each
of these populations after the end of study, similar methods from other
coefficients can be used.

Note that in both cases an initial number of infected individuals at
year 2021 is required ($I_{i,t=2021}$).

## Statistical models

### Predictive model of bat community composition change

The model provides some predictions for populations of different species
at different time points. We can then use this data to test whether
community dynamics are similar to declines in eastern North America
using a chi-2 goodness of fit test. Power analysis will also be
conducted. A single test should be conducted for all for species over
2021-2030 period. If the test finds significant difference, then an
exploratory test should be conducted (further details explained in next
sections) to determine new values for the coefficients using the data
from Saskatchewan bat community 2021-2030 using a genetic algorithm
optimization method. Then it should be determined if the sample size can
support the estimated effect size and again another round of power
analysis should be conducted.

## Transformations

N/A

## Inference criteria

For all analysis type I error threshold will be set to 0.05. Alpha error
correction will be done using the Bonferroni method to address the
effect of conducting multiple tests\\simultaneously on type I error.

## Data exclusion

See missing data.

## Missing data

We might expect bat records not to have a species identifier (e.g. not
able to identify species in field, etc). As this is data of interest
these records will have to be removed as there is no easy way to
re-examine an individual once released.

# References

## 

```{=tex}
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
```
\noindent
