---
title           : "Preregistration: Effects of white-nose syndrome (WNS) on bat community structure in Cypress Hills, Saskatchewan, Canada."
shorttitle      : "PreReg: Bat Community Structure in Cypress"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Sophia Fan
    affiliation : 1
  - name        : Keerthi Krutha
    affiliation : 3
  - name        : Javad Meghrazi
    affiliation : 1
  - name        : Tanya Strydom
    affiliation : 2
  - name        : Hannah Wilson
    affiliation : 3
      
affiliation:
    
  - id          : 1
    institution : University of British Columbia
  - id          : 2
    institution : University of Regina
  - id          : 3
    institution : Université de Montréal

output: prereg::cos_prereg
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

# Study Information

## Title

`r rmarkdown::metadata$title`

## Description

White-nose syndrome (WNS) is an infectious fungal disease caused by
*Pseudogymnoascus destructans* targeting hibernating bats. First
detected in New York City in the winter of 2006, it has since affected
and killed more than 6 million hibernating bats in eastern North America
@jachowski2014. Notably, the disease has had a pronounced effect on the
little brown bat, *Myotis lucifugus*, where it's predicted that regional
population collapse and extirpation could occur as soon as within the
next 16 years @frick2010a. Other bat species responded variably: some
studies show that silver-haired bat (*Lasionycteris noctivagans)*
detection rates increased following WNS [@jachowski2014a; @nocera2019],
while results for big brown bats (*Eptesicus fuscus*) are mixed, but
most studies seem to show that their numbers increased following
competitive relaxation from other declining bat species
[@jachowski2014b; @francl2012; @morningstar2019; @nocera2019a]. Because
WNS affects bat species differentially, community-level monitoring
studies are needed to fully appreciate and mitigate consequences for
community structure and ecosystem function @hoyt2021. In 2021, the first
instance of *P. destructans* with no bat mortality was reported in
Saskatchewan, Canada (Wilson pers. comm., Lausen-WCS Canada pers.
comm.), indicating westward disease spread into central Canada. This
study aims to inform local bat conservation efforts by compiling and
comparing bat community data from before and after the introduction of
*P. destructans* in the Cypress Hills, in addition to constructing
population projection models to predict species survivorship into the
future.

## Hypotheses

We expect to see white-nose syndrome affect bat species differently in
the Cypress Hill region of Saskatchewan, Canada (See Figure 1), in line
with similar trends found in affected eastern North American bat
populations. Specifically:

**Directional**: We expect to see pronounced population **declines** in
little brown bats (*Myotis lucifugus*).

**Directional**: We expect to see population **increases** in
silver-haired bats (*Lasionycteris noctivagans*) and big brown bats
(*Eptesicus fuscus*) due to reduced interspecific competition.

# Design Plan

<!-- In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration. -->

## Study type

**Observational Study**. Data is collected from study subjects that are
not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.

## Blinding

No blinding is involved in this study.

## Study design

**Design:** Case-control study design [@mann2003]. The study group will
be the bat community in Saskatchewan. Our control group would be the bat
community in the Cypress Hills before the detection of *P. destructans.*

**Assumption(s):** No two same individual bats were sampled before and
after *P. destructans* detection. Proportion of each species captured
are equal to their proportion in nature. Other factors that may impact
species decline are minimal.

## Randomization

Not applicable

# Sampling Plan

<!-- In this section we’ll ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study. -->

## Existing data

<!-- Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io). -->

**Registration prior to analysis of the data**. As of the date of
submission, the data exist and you have accessed it, though no analysis
has been conducted related to the research plan (including calculation
of summary statistics). A common situation for this scenario when a
large dataset exists that is used for many different studies over time,
or when a data set is randomly split into a sample for exploratory
analyses, and the other section of data is reserved for later
confirmatory data analysis.

## Explanation of existing data

```{=html}
<!-- If you indicate that you will be using some data that already exist in this study, please describe the steps you have taken to assure that you are unaware of any patterns or summary statistics in the data. This may include an explanation of how access to the data has been limited, who has observed the data, or how you have avoided observing any analysis of the specific data you will use in your study.

An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point.

Example: An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point. -->
```
Our dataset will be from 2010-2030, and the dataset from 2010-2021
already exists. Our access to the dataset has been limited because, all
team members except Ms. Hannah Wilson come from other departments and
laboratories with no prior studies conducted on bats. Ms. Wilson is a
new student to the bat research group to the University of Regina and
started in September 2021. The 2010-2021 dataset we plan to use for our
study has been collected by multiple trained researchers who are not
connected with our project. Ms. Wilson has seen the dataset with data
from 2019 but has not conducted any form of statistical analysis with
it. Also, as authors of this study, none of us are aware of any patterns
or summary statistics relating to the effect of a fungal disease on bats
of this dataset even though there are publications/results from previous
studies conducted by other researchers. We plan to only conduct any
detailed analysis with the data only after 2030 once all data collection
is complete.

## Data collection procedures

```{=html}
<!-- Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that donÍt include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

The answer to this question requires a specific set of instructions so that another person could repeat the data collection procedures and recreate the study population. Alternatively, if the study population would be unable to be reproduced because it relies on a specific set of circumstances unlikely to be recreated (e.g., a community of people from a specific time and location), the criteria and methods for creating the group and the rationale for this unique set of subjects should be clear.

Example: Participants will be recruited through advertisements at local pastry shops. Participants will be paid $10 for agreeing to participate (raised to $30 if our sample size is not reached within 15 days of beginning recruitment). Participants must be at least 18 years old and be able to eat the ingredients of the pastries. -->
```
**Study site:** West Block, Cypress Hills Provincial Park, Saskatchewan,
Canada.

**Study locations:** Previously established capture locations for bats
along Battle Creek, established by the University of Regina (Wilson
pers. comm.).

**Sampling period:** Mid-June to mid-August every year. Samples will be
collected until August 2030. Data points from this same period will be
used from records prior to *P. destructans* incidence starting from
August 2010.

**Data collection methods:** We will set up 1-3 mist nets at one site
each night. Nets will be set up across Battle Creek. Nets will be at
least 1m tall, and wide enough the cover the width of the creek. We will
raise nets for three hours after last light and check the nets every
10min in the first hour after last light and every 15min in the
subsequent two hours. We will capture and handle bats with
nitrile-gloves. Once a bat is caught it will be moved away from the net
in order to avoid attracting other individuals to the captured bat's
distress call. We will record the age, sex, mass and species of the bat.

In order to attribute the possible difference in bat population to the
fungal disease we will collect samples of *P. destructans* by swabbing
wings and muzzle of the bat follwing [@zukal2016] We will analyse swabs
using qPCR [@zukal2016].

We will also measure ambient temperature, cloud cover and wind speed
every night at each sampling site. In case of rain or if wind speed is
above 20km/s, we will stop sampling. When recording a capture, we will
record the date of sampling in the ISO format along with a three-letter
code of the sampling location, the genus and species of bat being
captured and initials of the personnel collecting the capture

## Sample size

```{=html}
<!-- Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis? For some studies, this will simply be the number of samples or the number of clusters. For others, this could be an expected range, minimum, or maximum number.

Example: Our target sample size is 280 participants. We will attempt to recruit up to 320, assuming that not all will complete the total task. -->
```
Our target sample size is 2000 individuals. However, we do not have a
pre-determined count and cannot control the number of bats we capture
each year.

## Sample size rationale

```{=html}
<!-- This could include a power analysis or an arbitrary constraint such as time, money, or personnel. This gives you an opportunity to specifically state how the sample size will be determined. A wide range of possible answers is acceptable; remember that transparency is more important than principled justifications. If you state any reason for a sample size upfront, it is better than stating no reason and leaving the reader to "fill in the blanks." Acceptable rationales include: a power analysis, an arbitrary number of subjects, or a number based on time or monetary constraints.

Example: We used the software program G*Power to conduct a power analysis. Our goal was to obtain .95 power to detect a medium effect size of .25 at the standard .05 alpha error probability. -->
```
We do not have control over how many bats will be captured in a night.
From anecdotal knowledge and previous experience we estimate that we
will to capture anywhere between 100-200 bats per summer. Therefore,
after twenty years of data collection we will have collected around 2000
bats.

## Stopping rule

```{=html}
<!-- If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection. 

You may specify a stopping rule based on p-values only in the specific case of sequential analyses with pre-specified checkpoints, alphas levels, and stopping rules. Unacceptable rationales include stopping based on p-values if checkpoints and stopping rules are not specified. If you have control over your sample size, then including a stopping rule is not necessary, though it must be clear in this question or a previous question how an exact sample size is attained.

Example: We will post participant sign-up slots by week on the preceding Friday night, with 20 spots posted per week. We will post 20 new slots each week if, on that Friday night, we are below 320 participants. -->
```
The station where researchers live during the field season is owned by
the University of Regina, who use the station for an undergraduate class
during the last week of August. The field station is closed after that,
so data collection will stop mid-August.

The project overall will end in 2030. *P. destructans* was first
detected in Saskatchewan in 2021. Ending the project 10 years after the
fungus was first detected allows us to observe bat community structure
at the start of the disease, directly after it should have killed the
largest number of bats, and ten years later when it should have become
endemic to the population.

# Variables

<!-- In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables which you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so. -->

## Manipulated variables

```{=html}
<!-- Describe all variables you plan to manipulate and the levels or treatment arms of each variable. This is not applicable to any observational study. For any experimental manipulation, you should give a precise definition of each manipulated variable. This must include a precise description of the levels at which each variable will be set, or a specific definition for each categorical treatment. For example, “loud or quiet,” should instead give either a precise decibel level or a means of recreating each level. 'Presence/absence' or 'positive/negative' is an acceptable description if the variable is precisely described.

Example: We manipulated the percentage of sugar by mass added to brownies. The four levels of this categorical variable are: 15%, 20%, 25%, or 40% cane sugar by mass. -->
```
'Raw' abundance records will be transformed to a proportional abundance
which will be used in the modelling process.

## Measured variables

```{=html}
<!-- Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

Observational studies and meta-analyses will include only measured variables. As with the previous questions, the answers here must be precise. For example, 'intelligence,' 'accuracy,' 'aggression,' and 'color' are too vague. Acceptable alternatives could be 'IQ as measured by Wechsler Adult Intelligence Scale' 'percent correct,' 'number of threat displays,' and 'percent reflectance at 400 nm.'

Example: The single outcome variable will be the perceived tastiness of the single brownie each participant will eat. We will measure this by asking participants ‘How much did you enjoy eating the brownie’ (on a scale of 1-7, 1 being 'not at all', 7 being 'a great deal') and 'How good did the brownie taste' (on a scale of 1-7, 1 being 'very bad', 7 being 'very good'). -->
```
We are going to be measuring the population size of the bat species in
the Cypress Hills before and after the detection of *P. destructans*. We
will measure this by counting how many bats of each species we catch in
mist nets each year.

## Indices

Not applicable.

# Analysis Plan

### Comparing population size of species before and after WNS

In order to compare population sizes before and after WNS, we would use
population size estimations at 2021 and 2030. \#\#\# Predictive model of
bat community composition change In order to investigate effect of WNS
on bat community as a whole instead of investigating species
individually, we have also developed a mathematical framework. Following
equations aim to describe changes in the size of populations of
different bat species. The aim of the model is to test if the effect of
WNS on bat community in Saskatchewan is similar to east coast
communities. Some coefficients should be estimated based on Saskatchewan
bat community data from 1990-2021 (growth rate, carrying capacity and
competition coefficients) and some from previous studies in east
coast(transmission rate)populations and then predication can be made for
Saskatchewan bat community. $$
N_{i,t+1} = N_{i,t}r_i (1+\frac{K_i-N_{i,t}- \sum_{i = 1}^{n}{\alpha_{i,j} N_{j,t}}}{K})-I_{i,t}
$$ Where: $N_{i,t}$ is the population size of species i at time t $r_i$
is the intrinsic growth rate of species i $K_i$ is the carrying capacity
for species i $\alpha_{i,j}$ is the coefficent for competition effect of
species j on species i. $I_{i,t}$ is the number of individuals infected
with WNS at time t. It's assumed that all of these individuals will dye
before next time point. This parameter is calculated as follow: $$
I_{i,t+1}=d_iI_{i,t}\frac{N_{i,t}}{k_i}
$$ Where $d_i$ is the transmission rate of WNS in species i. It's
assumed that transmission of diseas is density dependent.

Estimation of coefficients: $r$, $K$ and $\alpha$: They should be
calculated based on the data gathered sofar (1991-2021) using genetic
algorithm (or any other optimization method?).

$d$: depends on the objective of study: 1- If we are going to make
predictions for future dynamics of populations, previous studies should
be used to estimate this parameter. It should be considered that $d$
probably won't stay constant over time and evolutionary dynamics of host
and parasite would affect it. 2- If we are going to estimate $d$ in each
of these populations after the end of study, similar method as other
coefficients can be used.

Note that in both cases an initial number of infected individuals at
year 2021 is required ($I_{i,t=2021}$).

## Statistical models

### Comparing population size of species before and after WNS

Unpaired student t-tests will be used to compare population size
estimates at 2021 and 2023. The test should be one tail. Also power
analysis should be conducted using `pwr` package in R. \#\#\# Predictive
model of bat community composition change The model provides some
predictions for populations of different species at different time
points and we can test our hypothesis using a chi-2 goodness of fit
test. Besides, power analysis will be conducted. A single test should be
conducted for all for species over 2021-2030 period. If the test finds
significant difference, then an exploratory test should be conducted
(further details explained in next sections) to determine new values for
the coefficients using the data from Saskatchewan bat community
2021-2030 using a genetic algorithm optimization method. Then it should
be determined if the sample size can support the estimated effect size
and again another round of power analysis should be conducted.

## Transformations

```{=html}
<!-- If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process. If any categorical predictors are included in a regression, indicate how those variables will be coded (e.g. dummy coding, summation coding, etc.) and what the reference category will be.

Example: The "Effect of sugar on brownie tastiness" does not require any additional transformations. However, if it were using a regression analysis and each level of sweet had been categorically described (e.g. not sweet, somewhat sweet, sweet, and very sweet), 'sweet' could be dummy coded with 'not sweet' as the reference category. -->
```
N/A

## Inference criteria

```{=html}
<!-- What criteria will you use to make inferences? Please describe the information youÍll use (e.g. p-values, bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

p-values, confidence intervals, and effect sizes are standard means for making an inference, and any level is acceptable, though some criteria must be specified in this or previous fields. Bayesian analyses should specify a Bayes factor or a credible interval. If you are selecting models, then how will you determine the relative quality of each? In regards to multiple comparisons, this is a question with few "wrong" answers. In other words, transparency is more important than any specific method of controlling the false discovery rate or false error rate. One may state an intention to report all tests conducted or one may conduct a specific correction procedure; either strategy is acceptable.

Example: We will use the standard p<.05 criteria for determining if the ANOVA and the post hoc test suggest that the results are significantly different from those expected if the null hypothesis were correct. The post-hoc Tukey-Kramer test adjusts for multiple comparisons. -->

For all analysis type I error treshold will be set to 0.05. Alpha error correction 
will be done using Bonferroni method to address effect of conducting multiple tests\
simultaneously on type I error. 
```
## Data exclusion

```{=html}
<!-- How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled? Will you use any awareness check? Any rule for excluding a particular set of data is acceptable. One may describe rules for excluding a participant or for identifying outlier data.

Example: No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the three tastiness indices. Outliers will be included in the analysis. -->
```
See missing data

## Missing data

We might expect bat records not to have a species identifier. As this is
the data of interest these records will have to be removed as there is
no feasible way to impute the missing data.

## Exploratory analyses (optional)

If the test finds significant difference, then an exploratory test
should be conducted to determine new values for the coefficients using
the data from Saskatchewan bat community 2021-2030 using a genetic
algorithm optimization method. Then it should be determined if the
sample size can support the estimated effect size and again another
round of power analysis should be conducted.

# Other

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("data/sim_data.R")
```

```{r plot, echo=FALSE, fig.cap="Expected changes in bat species frequencies after the introduction of white nose syndrome in 2019 the Cypress hill region of Saskatchewan Province.", message=FALSE, warning=FALSE}
p
```

# References

## 

```{=tex}
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
```
\noindent
