---
title           : "Preregistration: Bat Community Structure In the Cypress Hills pre/post White-Nose Syndrome"
shorttitle      : "PreReg: Bat Community Structure in Cypress"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Sophia Fan
    affiliation : 1
  - name        : Keerthi Krutha
    affiliation : 3
  - name        : Javad Meghrazi
    affiliation : 1
  - name        : Tanya Strydom
    affiliation : 2
  - name        : Hannah Wilson
    affiliation : 3
      
affiliation:
    
  - id          : 1
    institution : University of British Columbia
  - id          : 2
    institution : University of Regina
  - id          : 3
    institution : Université de Montréal

output: prereg::cos_prereg
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

# Study Information

## Title

`r rmarkdown::metadata$title` Effects of white-nose syndrome (WNS) on
bat community structure in Cypress Hills, Saskatchewan, Canada.

## Description

White-nose syndrome (WNS) is an infectious fungal disease caused by
*Pseudogymnoascus destructans* targeting hibernating bats. First
detected in New York City in the winter of 2006, it has since affected
and killed more than 6 million hibernating bats in eastern North America
@jachowski2014. Notably, the disease has had a pronounced effect on the
little brown bat, *Myotis lucifugus*, where it's predicted that regional
population collapse and extirpation could occur as soon as within the
next 16 years @frick2010a. Other bat species responded variably: some
studies show that silver-haired bat (*Lasionycteris noctivagans)*
detection rates increased following WNS [@jachowski2014a; @nocera2019],
while results for big brown bats (*Eptesicus fuscus*) are mixed, but
most studies seem to show that their numbers increased following
competitive relaxation from other declining bat species
[@jachowski2014b; @francl2012; @morningstar2019; @nocera2019a]. Because
WNS affects bat species differentially, community-level monitoring
studies are needed to fully appreciate and mitigate consequences for
community structure and ecosystem function @hoyt2021. In 2021, the first
instance of *P. destructans* with no bat mortality was reported from the
Cypress Hills Interprovincial Park in Saskatchewan, Canada (Wilson pers.
comm., Lausen-WCS Canada pers. comm.), indicating westward disease
spread into central Canada. This study aims to inform local bat
conservation efforts by compiling and comparing bat community data from
before and after the introduction of WNS in Cypress Hills, in addition
to constructing population projection models under WNS to predict
species survivorship into the future.

## Hypotheses

We expect to see white-nose syndrome affect bat species differently in
the Cypress Hill region of Saskatchewan, Canada (See Figure 1).

**Directional**: Like other parts of eastern North America, we expect to
see pronounced population **declines** in little brown bats (*Myotis
lucifugus*).

**Directional**: We expect to see population **increases** in
silver-haired bats (*Lasionycteris noctivagans*) and big brown bats
(*Eptesicus fuscus*) due to reduced interspecific competition. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("data/sim_data.R")
```

```{r plot, echo=FALSE, fig.cap = "Expected changes in bat species frequencies after the introduction of white nose syndrome in 2019 the Cypress hill region of Saskatchewan Province.", message=FALSE, warning=FALSE}
p
```

# Design Plan

<!-- In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration. -->

## Study type

**Observational Study**. Data is collected from study subjects that are
not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.

## Blinding

No blinding is involved in this study.

## Study design

Timing: The Bat Community in the Cypress Hills has been monitored since
1991, and we will continue to monitor the community until 2030.
Monitoring will start around mid-June every year, depending upon permit
permissions. Monitoring will finish in mid-August.

Study sites: Researchers will place mist nets at various points along
Battle Creek in the Cypress Hills West Block. We will choose these sites
based on records from previous capture data. We will continue to return
to these sites each year until 2030. We will also attempt to net at each
site at similar dates each year.

Field methods: We will raise 1-3 mist nets across Battle Creek at each
net site. Number of nets will be chosen based on personnel, time, and
previous net set-ups. Each mist net will be raised at last light and we
will monitor net activity for three hours per night. Every hour we will
record ambient temperature, cloud cover and wind speed. We will close
nets if it is raining or wind speed is above 20 km/s.

We will check the nets every 10 minutes for one hour after last light,
and every 15 minutes for the two proceeding hours. If a bat is captured
it will be extracted and moved away from the mist nets in order to
reduce other bats being attracted to the distress call of the captured
bat. We will take data on each captured bats species, age, sex, and
mass. Bats will be released at most half an hour after capture.

## Randomization

Not applicable

# Sampling Plan

<!-- In this section we’ll ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study. -->

## Existing data

<!-- Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io). -->

**Registration prior to analysis of the data**. As of the date of
submission, the data exist and you have accessed it, though no analysis
has been conducted related to the research plan (including calculation
of summary statistics). A common situation for this scenario when a
large dataset exists that is used for many different studies over time,
or when a data set is randomly split into a sample for exploratory
analyses, and the other section of data is reserved for later
confirmatory data analysis.

## Explanation of existing data

```{=html}
<!-- If you indicate that you will be using some data that already exist in this study, please describe the steps you have taken to assure that you are unaware of any patterns or summary statistics in the data. This may include an explanation of how access to the data has been limited, who has observed the data, or how you have avoided observing any analysis of the specific data you will use in your study.

An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point.

Example: An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point. -->
```
Capture data from bats in the Cypress Hills has been collected since
1991, and so we have data for this project from 1991 until now. The
existing data was collected by other researchers and was used for many
different studies over time. These researchers left their raw data
available for others to use. Currently, only one of the project members
has seen the original data. However, they have only seen the data from
1991-1992, and the 2019 dataset and they have not conducted any prior
analyses or investigations of it. The other project members have not
seen any of the existing data yet. Furthermore, this project assesses
community structure changes before and after the detection of White-nose
syndrome in Saskatchewan until 2030. We will not analyze the data until
after 2030.

## Data collection procedures

```{=html}
<!-- Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that donÍt include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

The answer to this question requires a specific set of instructions so that another person could repeat the data collection procedures and recreate the study population. Alternatively, if the study population would be unable to be reproduced because it relies on a specific set of circumstances unlikely to be recreated (e.g., a community of people from a specific time and location), the criteria and methods for creating the group and the rationale for this unique set of subjects should be clear.

Example: Participants will be recruited through advertisements at local pastry shops. Participants will be paid $10 for agreeing to participate (raised to $30 if our sample size is not reached within 15 days of beginning recruitment). Participants must be at least 18 years old and be able to eat the ingredients of the pastries. -->
```
Monitoring will start around mid-June every year, depending upon permit
permissions. Monitoring will finish in mid-August. We will monitor
community structure using the capture data we collect from mist nets.

We will set up 1-3 mist nets at one site each night. Nets will be set up
across Battle Creek. Nets will be at least 1m tall, and wide enough the
cover the width of the creek. Nets will be set up at transition sites
where the flyway above Battle Creek transitions from a cluttered
environment with lots of debris to open sky. Nets will also be placed
over calm sections of water, without ripples.

Each mist net will be raised at last light and we will monitor net
activity for three hours per night. Every hour we will record ambient
temperature, cloud cover and wind speed. We will close nets if it is
raining or wind speed is above 20 km/s.

We will check the nets every 10 minutes for one hour after last light,
and every 15 minutes for the two proceeding hours. If a bat is captured
it will be extracted and moved away from the mist nets in order to
reduce other bats being attracted to the distress call of the captured
bat. We will take data on each captured bats species, age, sex, and
mass. Bats will be released at most half an hour after capture.

We will continue to return to these sites each year until 2030. We will
also attempt to net at each site at similar dates each year.

## Sample size

```{=html}
<!-- Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis? For some studies, this will simply be the number of samples or the number of clusters. For others, this could be an expected range, minimum, or maximum number.

Example: Our target sample size is 280 participants. We will attempt to recruit up to 320, assuming that not all will complete the total task. -->
```
Our target sample size is 4000 individuals. However, we do not have a
pre-determined count and cannot control the number of bats we capture
each year.

## Sample size rationale

```{=html}
<!-- This could include a power analysis or an arbitrary constraint such as time, money, or personnel. This gives you an opportunity to specifically state how the sample size will be determined. A wide range of possible answers is acceptable; remember that transparency is more important than principled justifications. If you state any reason for a sample size upfront, it is better than stating no reason and leaving the reader to "fill in the blanks." Acceptable rationales include: a power analysis, an arbitrary number of subjects, or a number based on time or monetary constraints.

Example: We used the software program G*Power to conduct a power analysis. Our goal was to obtain .95 power to detect a medium effect size of .25 at the standard .05 alpha error probability. -->
```
We do not have control over how many bats we will capture a night. We
are assuming based anecdotal knowledge and previous experience that in
three months we will capture 100-200 bats, and after forty years of data
collection we will collect around 4000 bats.

## Stopping rule

```{=html}
<!-- If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection. 

You may specify a stopping rule based on p-values only in the specific case of sequential analyses with pre-specified checkpoints, alphas levels, and stopping rules. Unacceptable rationales include stopping based on p-values if checkpoints and stopping rules are not specified. If you have control over your sample size, then including a stopping rule is not necessary, though it must be clear in this question or a previous question how an exact sample size is attained.

Example: We will post participant sign-up slots by week on the preceding Friday night, with 20 spots posted per week. We will post 20 new slots each week if, on that Friday night, we are below 320 participants. -->
```
Data collection will end each year in mid-August. The field station
where researchers live during the field season is owned by the
University of Regina, who use the station for an undergraduate class
during the last week of August. The field station is closed after that,
so data collection will stop mid-August

The project overall will end in 2030. White-Nose Syndrome was first
detected in Saskatchewan in 2021. Ending the project 10 years after WNS
was first detected allows us to observe bat community structure at the
start of the disease, directly after it should have killed the largest
number of bats, and ten years later when it should have become endemic
to the population.

# Variables

<!-- In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables which you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so. -->

## Manipulated variables

```{=html}
<!-- Describe all variables you plan to manipulate and the levels or treatment arms of each variable. This is not applicable to any observational study. For any experimental manipulation, you should give a precise definition of each manipulated variable. This must include a precise description of the levels at which each variable will be set, or a specific definition for each categorical treatment. For example, “loud or quiet,” should instead give either a precise decibel level or a means of recreating each level. 'Presence/absence' or 'positive/negative' is an acceptable description if the variable is precisely described.

Example: We manipulated the percentage of sugar by mass added to brownies. The four levels of this categorical variable are: 15%, 20%, 25%, or 40% cane sugar by mass. -->
```
*TK* I'm assuming 'raw' abundance records will be transformed to some
measure of frequency/ratio or otherwise 'extrapolated' to some measure
of abundance?

## Measured variables

```{=html}
<!-- Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

Observational studies and meta-analyses will include only measured variables. As with the previous questions, the answers here must be precise. For example, 'intelligence,' 'accuracy,' 'aggression,' and 'color' are too vague. Acceptable alternatives could be 'IQ as measured by Wechsler Adult Intelligence Scale' 'percent correct,' 'number of threat displays,' and 'percent reflectance at 400 nm.'

Example: The single outcome variable will be the perceived tastiness of the single brownie each participant will eat. We will measure this by asking participants ‘How much did you enjoy eating the brownie’ (on a scale of 1-7, 1 being 'not at all', 7 being 'a great deal') and 'How good did the brownie taste' (on a scale of 1-7, 1 being 'very bad', 7 being 'very good'). -->
```
We are going to be measuring the population size of the bat species in
the Cypress Hills before and after the detection of White-nose Syndrome.
We will measure this by counting how many bats of each we catch in mist
nets each year.

## Indices

```{=html}
<!-- If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

If you are using multiple pieces of data to construct a single variable, how will this occur? Both the data that are included and the formula or weights for each measure must be specified. Standard summary statistics, such as "means" do not require a formula, though more complicated indices require either the exact formula or, if it is an established index in the field, the index must be unambiguously defined. For example, "biodiversity index" is too broad, whereas "Shannon’s biodiversity index" is appropriate.

Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->
```
Enter your response here.

# Analysis Plan

```{=html}
<!-- You may describe one or more confirmatory analysis in this preregistration. Please remember that all analyses specified below must be reported in the final article, and any additional analyses must be noted as exploratory or hypothesis generating.

A confirmatory analysis plan must state up front which variables are predictors (independent) and which are the outcomes (dependent), otherwise it is an exploratory analysis. You are allowed to describe any exploratory work here, but a clear confirmatory analysis is required. -->
```
## Statistical models

```{=html}
<!-- What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions, subgroup analyses, pairwise or complex contrasts, or follow-up tests from omnibus tests. If you plan on using any positive controls, negative controls, or manipulation checks you may mention that here. Remember that any test not included here must be noted as an exploratory test in your final article.

This is perhaps the most important and most complicated question within the preregistration. As with all of the other questions, the key is to provide a specific recipe for analyzing the collected data. Ask yourself: is enough detail provided to run the same analysis again with the information provided by the user? Be aware for instances where the statistical models appear specific, but actually leave openings for the precise test. See the following examples:

- If someone specifies a 2x3 ANOVA with both factors within subjects, there is still flexibility with the various types of ANOVAs that could be run. Either a repeated measures ANOVA (RMANOVA) or a multivariate ANOVA (MANOVA) could be used for that design, which are two different tests. 
- If you are going to perform a sequential analysis and check after 50, 100, and 150 samples, you must also specify the p-values you’ll test against at those three points.

Example:  We will use a one-way between subjects ANOVA to analyze our results. The manipulated, categorical independent variable is 'sugar' whereas the dependent variable is our taste index. -->
```
Enter your response here.

## Transformations

```{=html}
<!-- If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process. If any categorical predictors are included in a regression, indicate how those variables will be coded (e.g. dummy coding, summation coding, etc.) and what the reference category will be.

Example: The "Effect of sugar on brownie tastiness" does not require any additional transformations. However, if it were using a regression analysis and each level of sweet had been categorically described (e.g. not sweet, somewhat sweet, sweet, and very sweet), 'sweet' could be dummy coded with 'not sweet' as the reference category. -->
```
Enter your response here.

## Inference criteria

```{=html}
<!-- What criteria will you use to make inferences? Please describe the information youÍll use (e.g. p-values, bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

p-values, confidence intervals, and effect sizes are standard means for making an inference, and any level is acceptable, though some criteria must be specified in this or previous fields. Bayesian analyses should specify a Bayes factor or a credible interval. If you are selecting models, then how will you determine the relative quality of each? In regards to multiple comparisons, this is a question with few "wrong" answers. In other words, transparency is more important than any specific method of controlling the false discovery rate or false error rate. One may state an intention to report all tests conducted or one may conduct a specific correction procedure; either strategy is acceptable.

Example: We will use the standard p<.05 criteria for determining if the ANOVA and the post hoc test suggest that the results are significantly different from those expected if the null hypothesis were correct. The post-hoc Tukey-Kramer test adjusts for multiple comparisons. -->
```
## Data exclusion

```{=html}
<!-- How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled? Will you use any awareness check? Any rule for excluding a particular set of data is acceptable. One may describe rules for excluding a participant or for identifying outlier data.

Example: No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the three tastiness indices. Outliers will be included in the analysis. -->
```
Enter your response here.

## Missing data

We might expect bat records not to have a species identifier. As this is
the data of interest these records will have to be removed as there is
no feasible way to impute the missing data.

## Exploratory analyses (optional)

```{=html}
<!-- If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

Example: We expect that certain demographic traits may be related to taste preferences. Therefore, we will look for relationships between demographic variables (age, gender, income, and marital status) and the primary outcome measures of taste preferences. -->
```
Enter your response here.

# Other

## Other (Optional)

<!-- If there is any additional information that you feel needs to be included in your preregistration, please enter it here. Literature cited, disclosures of any related work such as replications or work that uses the same data, or other context that will be helpful for future readers would be appropriate here. -->

Enter your response here.

# References

## 

```{=tex}
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
```
\noindent
